On the presentation of orthogonal groups over polynomial rings

### 2019-06-17 dl-paper.tex
Minor fixes

\subsection{$\K_2$-groups and symbols}

In our computations we use two families of explicit elements of $\K_2(\Phi, R)$ called {\it Steinberg and Dennis--Stein symbols}. Notice that our treatment of symbols follows~\cite{DS73}. Recall that Steinberg symbols are defined for arbitrary $s, t \in R^\times$ as follows:
\begin{equation} \label{eq:steinberg} \{ s, t \}_\alpha = h_\alpha(st) \cdot h_\alpha^{-1}(s) \cdot h_\alpha^{-1}(t). \end{equation}
In turn, Dennis--Stein symbols are defined for arbitrary $a, b\in R$ satisfying $1 + ab \in R^\times$:
\begin{equation} \label{eq:dennis-stein}  \langle a,b \rangle _ \alpha = x_{-\alpha}\left(\tfrac{- b}{1 + ab}\right) \cdot x_{\alpha}(a) \cdot x_{-\alpha}(b) \cdot x_{\alpha}\left(\tfrac{- a}{1+ab}\right) \cdot h_{\alpha}^{-1}(1 + ab). \end{equation} 
Dennis--Stein symbol $\langle a, b \rangle_\alpha$ can be expressed through Steinberg symbols in the special case when either $a$ or $b$ is an invertible element of $R$. More specifically, the following formulae hold (cf.~\cite[p.~250]{DS73}).
\begin{equation} \label{DS-S-relationship} \langle a, b \rangle_\alpha = \{-a, 1+ab\}_\alpha\text{ for } a, 1+ab\in R^\times,\ \
 \{ s, t \}_\alpha = \left\langle -s, \tfrac{1 - t}{s} \right\rangle_\alpha\text{ for } s, t\in R^\times. \end{equation}
%Using~\eqref{eq:conj-h-h} it is easy to verify that \[[h_\alpha(s), h_\beta(t)] = \{s, t^{\langle \alpha, \beta \rangle}\}_\alpha = \{t, s^{\langle \beta, \alpha \rangle}\}_\beta^{-1}.\] 
Steinberg and Dennis--Stein symbols depend only on the length of $\alpha$, in particular they do not depend on $\alpha$ if $\Phi$ is simply-laced. If $\Phi$ happens to be {\it nonsymplectic}, i.\,e. $\Phi \neq \rA_1, \rB_2, \rC_{\geq 3}$, Steinberg symbols are antisymmetric and bimultiplicative, i.\,e. they satisfy the following identities: \begin{equation} \label{eq:symbol-properties} \{ u, st \} = \{ u, s\} \{ u, t \}, \ \{ u, v \} = \{ v, u\}^{-1}. \end{equation}
For these and other properties of symbols we refer the reader to~\cite{DS73}.

@@

\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate}
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} $S$ preserves coequalizers of kernel pairs.
\end{enumerate}

@@

%TODO: Give the definition of $D_{R, I}$ and of the category of Pairs
\begin{prop}\label{characterization}
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is the universal relative central extension of the map $S(R) \to S(R/I)$.
       Denote the kernel of $\mu$ by $C(R, I)$.
       Then $C(R, I)$ is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is a natural exact sequence:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & C(R, I).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from \cite[Chapter~IV]{Kbook} (cf. Exercises~1.8--1.9 ibid, here we use assumptions~\eqref{req:subfunc}--\eqref{req:uce}). %provided one verifies that all the intermediate isomorphisms are natural in $R$.

 Let us verify the second claim. From~\eqref{req:left-exact} and $\varphi(G_i) \subseteq \Ker(G(p_i))$ we conclude that $G_0 \cap G_1 \subseteq \Ker(\varphi \colon S(D_{R, I}) \to G(D_{R, I}))$
  therefore $G_0 \cap G_1$ is a central subgroup of $S(D_{R, I})$.
 From~\eqref{req:uce} we also obtain that $H_1(S(D_{R,I})) = H_2(S(D_{R,I})) = 0$ .
 The last two assertions allow us to invoke~\cite[Proposition~6]{Lo78} and conclude that the extension $S(R, I) \to S(R)$ is a universal relative central extension of the coequalizer $\mathrm{coeq}(d_0, d_1)$
  in the sense of~\cite[\S~3]{Lo78}. 
 Notice that this coequalizer coincides with $S(R) \to S(R/I)$ by~\eqref{req:coeq}. 
 Finally, the kernel of $\mu$ is isomorphic to $\HH_3(S(R), S(R/I), \ZZ)$ by~\cite[Th{\'e}or{\`e}me~2]{Lo78}.

@@

The principal example of a triple $(G,P,S)$ satisfying the above conditions is as follows.
For $n \geq 5$ take the orthogonal group functor $\mathrm{O}_{2n}(-)$ and its elementary subfunctor $\mathrm{EO}_{2n}(-)$ as functors $G_n$ and $P_n$, respectively. Now set $S_n = \St(\rD_n, -)$. From [Stein's paper] and [Lemma on centrality] it follows that $S_n$ and $P_n$ satisfy~\eqref{req:uce}.
 The other requirements can be verified in a straightforward manner. We write $KO_i(n, R)$ as a shorthand for $\K_i(G_n, P_n, R)$.

@@

\section{First injectivity theorem}

If we denote the kernel of the natural transformation $\mu$ from the definition of $\St(\Phi, R, I)$ by $C(\Phi, R, I)$ we obtain a natural exact sequence:
\begin{equation} \label{relSteinberg-def}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r]^{j} & \St(\Phi, R, I) \ar[r]^{\mu} & \St(\Phi, R) \ar[r]^{\pi} & \St(\Phi, R/I) \ar[r] & 1.}
\end{equation}
The group $C(\Phi, R, I)$ is trivial when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits, see \cite[Lemma~8]{S15}.
The following result plays a key role in the proof of~\cref{thm41}.
\begin{lemma} \label{lem:prop41}
In the notation of~\cref{lem:lemma32} assume additionally that the residue field $k$ is of characteristic $\neq 2$.
Then the canonical map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective for $\ell \geq 7$.
\end{lemma}
\begin{proof}
  Consider the following commutative diagram with rows obtained from~\cref{characterization}:
\begin{equation*}\xymatrix{
 KO_3(2\ell, B) \ar[r] \ar[d] & KO_3(2\ell, k[X]) \ar[d]_{f'} \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
 KO_3(2\ell, R) \ar[r]        & KO_3(2\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
In view of~\cref{lem:Panin-stability} the central vertical map $f'$ is precisely the canonical map $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$.
Invoking~\cref{bass-ft} (or~\cite[Corollary~5.3]{Ho05}) we obtain that \[\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k) \oplus \GW_2^{[3]}(k).\]
By \cite[Lemma~2.2]{FRS12} the group $\GW_2^{[3]}(k)$ is trivial, therefore $f'$ (and hence $f$) are surjective.
\end{proof}

@@

\begin{comment}
\begin{rem}
We can take $\mathbb R^l$ with a base $\chi_1,\ldots,\chi_l$ and visualise root systems
\[\rD_l=\{\pm\chi_i\pm\chi_j\mid1\leq i\neq j\leq l\}\]
and $\rC_l=\rD_l\cup\{\pm2\chi_i\mid1\leq i\leq l\}$. Then we immediately see that $\alpha+\beta\in\rC_l\setminus\rD_l$ exactly means that either $\{\alpha,\beta\}$ or $\{-\alpha,-\beta\}$ equals $\{\chi_i-\chi_j,\chi_i+\chi_j\}$. Then for some $\sigma$ from the Weyl group of $\rD_l$ we get 
\[ \{\sigma\alpha,\sigma\beta\}=\{\chi_{l-1}-\chi_l, \chi_{l-1}+\chi_l\}=\{\alpha_{l-1},\alpha_l\} \]
(the horns of the Dynkin diagram $\rD_l$). Another condition on $\alpha$ and $\beta$ equivalent to $\alpha+\beta\in\rC_l\setminus\rD_l$ is that they do not lie in any subsystem of $\rD_l$ of type $\rA_m$ for $m\geq4$.
\end{rem}
Everywhere below we set $t = X^{-1}$.
\begin{lemma} \label{claim2} In $\St^{\leq n}(\Phi, B)$ relations $\text{R3}^\angle_{d, e}$ imply $\text{R3}^{\bot,0}_{d, e}$ for $1\leq d\leq e \leq n$. \end{lemma}
\begin{proof}
Without loss of generality we may assume that $\alpha, \gamma$ fit into some $\mathsf{A}_3$-triple $(\alpha, \beta, \gamma)$.
 \begin{align*}
 {}^{x_\gamma(t^d)}x_\alpha(bt^{e}) = {}^{x_\gamma(t^d)}[x_{-\beta}(b't^{e}), x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e,0}$}\\
 = [x_{-\beta}(b't^{e}), {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R3}^\angle_{d, e}$}\\
 = [[x_{-\beta-\gamma}(b''t^{e-d}), x_\gamma(t^d)], {}^{x_\gamma(t^d)}x_{\alpha+\beta}(1)] & \text{ by $\text{R2}_{e-d,d}$}\\
 = [x_{-\beta-\gamma}(b''t^{e-d)}), [x_\gamma(t^d), x_{\alpha+\beta}(1)]] & \text{ by~\eqref{eq:H1iii} and $\text{R3}^\bot_{e-d,0}$}\\
 = x_{\alpha}(b'''t^{e}) & \text{ by $\text{R2}_{0,d}$ and $\text{R2}_{e-d,d}$.} \end{align*}
Usual identities for structure constants imply (cf.~\cite[p.~12]{Re75}):
 $b'''=N_{-\beta-\gamma, \alpha+\beta+\gamma} \cdot N_{\gamma, \alpha+\beta} \cdot N_{-\beta-\gamma, \gamma} \cdot N_{-\beta, \alpha+\beta} \cdot b = b,$ from which the claim follows.
\end{proof}

\begin{lemma} \label{claim3}
 In $\St^{\leq n}(\Phi, B)$ relations $\bot_0(d, e)$, $\angle(d, d)$ and $\angle(e, e)$ imply $\bot(d, e)$ for $d\leq e \leq n$.
\end{lemma}
\begin{proof}
 \begin{align*} 
 x_{\alpha+\beta+\gamma}(abt^{e}) = {}^{x_{-\beta}(t^d)}\!x_{\alpha+\beta+\gamma}(abt^{e}) & \text{ by $\text{R3}^{\bot, 0}_{d, e}$}\\
 = {}^{x_{-\beta}(t^d)}\![x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$} \\
 = [x_{\alpha}(b't^{e}) \cdot x_{\alpha+\beta}(b't^{e-d}), x_\gamma(at^d)] & \text{ by $\text{R2}_{e-d,d}$ and $\text{R3}^\angle_{d, d}$ }\\
 = {}^{x_{\alpha}(b't^{e})}\!x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by~\eqref{eq:H1ii} and $\text{R2}_{e-d,d}$} \\ = x_{\alpha+\beta+\gamma}(abt^{e}) [x_{\alpha}(b't^{e}), x_\gamma(at^d)] & \text{ by $\text{R3}^\angle_{e, e}$. \qedhere}
\end{align*}
\end{proof}

We claim that $G_n\cong G_{n+1}$, more precisely, we construct an inverse homomorhism to the natural map $G_n\rightarrow G_{n+1}$. With this end we define elements $s_\alpha(a)\in G_n$ for $a\in R_{n+1}$ and prove the above relations for these elements. We start with the following easy fact.

\begin{lemma}
\label{welldef}
Take $d$, $e\leq n$, $d+e=n+1$, $a\in R_d$, $b\in R_e$, and consider an embedding $\iota\colon\rA_4\hookrightarrow\rD_l$. For the chosen simple roots $\alpha_k$ of $\rA_4$ denote $\beta_k=\iota(\alpha_k)$. Then holds
\[ [x_{\beta_1}(a),\,x_{\beta_2+\beta_3}(b)]=[x_{\beta_1+\beta_2}(Nabt\inv),\,x_{\beta_3}(t)], \]
where $N=N_{\beta_1,\beta_2}N_{\beta_2,\beta_3}$.
\end{lemma}
\begin{proof}
Observe that $d$, $e\geq1$. Take $x=x_{\beta_1}(a)$, $y=x_{\beta_2}(N_{\beta_3,\beta_2}bt\inv)$, $z=x_{\beta_3}(t)$. Since $x$ and $z$ lie in the subsystem of type $\rA_4$, we get by R3.1$\!\,_n$ that they commute. Then Hall--Witt identity gives us
\[ \,^y[x,\,[y\inv,\,z]]=\,^x[[x\inv,\,y],\,z], \]
where $[x,\,[y\inv,z]]=[x_{\beta_1}(a),\,x_{\beta_2+\beta_3}(b)]$ and $[[x\inv,y],\,z]=[x_{\beta_1+\beta_2}(Nabt\inv),\,x_{\beta_3}(t)]$ by R2$\!\,_n$. Here $x$ commutes with $x_{\beta_1+\beta_2}(Nabt\inv)$ and $x_{\beta_3}(t)$, and $y$ commutes with $x_{\beta_2+\beta_3}(b)$ by R3.1$\!\,_n$. Finally, $[x_{\beta_1}(-a),\,y]=x_{\beta_1+\beta_2}(Nabt\inv)$ also commutes with $x_{\beta_2+\beta_3}(b)$, what finishes the proof.
\end{proof}

The above lemma allows us for $a\in R_{n+1}$ and $\alpha\in\rD_l$ chose an arbitrary decomposition $\alpha=\beta+\gamma$ and define $s_\alpha(a)=[x_\beta(N_{\beta,\gamma}\,at\inv),\,x_\gamma(t)]$. Indeed, take another decomposition $\alpha=\beta'+\gamma'$, we may assume that subsystem $\Psi$ generated by $\beta$, $\beta'$, $\gamma$, $\gamma'$ has type $\rA_3$ If it is contained in the subsystem of type $\rA_4$, we are done by Lemma~\ref{welldef}. If not, we may assume $\Psi=\langle\alpha_{l-2},\alpha_{l-1},\alpha_l\rangle$ and since $l\geq5$ take two subsystems of type $\rA_4$, say, $\Psi_1$ and $\Psi_2$, which overlap by $\rA_3$, and $\beta$, $\gamma\in\Psi_1$, $\beta'$, $\gamma'\in\Psi_2$. Therefore, there exist $\beta''$, $\gamma''\in\Psi_1\cap\Psi_2$ with $\alpha=\beta''+\gamma''$ and we can again apply Lemma~\ref{welldef} to get the claim.

Moreover, with the assumption of Lemma~\ref{welldef} on $a$ and $b$ we have
\[ [x_{\beta}(a),\,x_{\gamma}(b)]=s_{\alpha}(N_{\beta,\gamma}\,ab). \]

The following observation is obvious.
\begin{lemma} \label{commute}
Consider an embedding $\iota\colon\rA_4\hookrightarrow\rD_l$ and denote $\beta_k=\iota(\alpha_k)$ for simple roots $\alpha_k$ of $\rA_4$. Then for any $a\in R_d$, $b\in R_e$, $c\in R_f$ with $d$, $e$, $f\leq n$ holds
\[ [[x_{\beta_1}(a),\,x_{\beta_2}(b)],\,x_{\beta_4}(c)]=1. \]
\end{lemma}

Consider now any two roots $\alpha$, $\beta\in\rD_l$ such that $\alpha+\beta\not\in\rC_l\cup0$. Since $l\geq5$, there exists a subsystem $\Psi$ of type $\rA_4$, containing $\alpha$ and $\beta$. Then we can take any $a\in R_{n+1}$ and $b\in R_d$ for $d\leq n$ and apply Lemma~\ref{commute} for $\Psi$ to get
\begin{equation} \label{sx} [s_\alpha(a),\,x_\beta(b)]=1. \end{equation}
Next, using this fact we can apply again Lemma~\ref{commute} for $\Psi$ and get
\[ [s_\alpha(a),\,s_\beta(c)]=1 \]
for any $c\in R_{n+1}$. Therefore, R3.1$\!\,_{n+1}$ holds for $s_\alpha$'s.

Now one can take $a\in R_d$ for $d\leq0$ and $b\in R_{n+1}$, and repeat the proof of Lemma~\ref{welldef} using~(\ref{sx}) to get
\[ [x_\beta(a),\,s_\gamma(b)]=\begin{cases}s_{\beta+\gamma}(N_{\beta\gamma}\,ab),&d=0,\\x_{\beta+\gamma}(N_{\beta\gamma}\,ab),&d<0.\end{cases} \]
This finishes the proof of R2$\!\,_{n+1}$.

The proof of R1$\!\,_{n+1}$ is left to the reader. The next lemma is devoted to R3.2$\!\,_{n+1}$.

\begin{lemma} For $a\in R_d$, $b\in R_e$, $d$, $e\leq n$, $d+e=n+1$ holds \[ [x_{\alpha_{l-1}}(a),\,x_{\alpha_l}(b)]=1. \] \end{lemma}
\begin{proof} Take $x=x_{\alpha_{l}-\alpha_{l-2}}(N_{\alpha_{l}-\alpha_{l-2},\alpha_{l-2}}bt\inv)$, $y=x_{\alpha_{l-1}-\alpha_{l-2}}(N_{\alpha_{l-2},\alpha_{l-1}-\alpha_{l-2}}at\inv)$, and $z=x_{\alpha_{l-2}}(t)$. Then
\[ [x_{\alpha_{l-1}}(a),\,x_{\alpha_l}(b)]=[[y\inv,\,z],\,[x,\,z]]=\,^{y\inv z}[y,\,[z\inv,\,x]]\cdot[y\inv,\,[x,\,z]]. \]
Next, by R3.2$\!\,_n$ we know that $x$ and $y$ commute, and therefore 
\[ \,^{y\inv z}[y,\,[z\inv,\,x]]=[[y\inv,\,z],\,x]=x_{\alpha_{l-1}+\alpha_{l}-\alpha_{l-2}}(Nabt\inv) \]
for $N=N_{\alpha_{l-1},\alpha_{l}-\alpha_{l-2}}N_{\alpha_{l}-\alpha_{l-2},\alpha_{l-2}}$. Finally,
\[ [y\inv,\,[x,\,z]]=x_{\alpha_{l-1}-\alpha_{l-2}+\alpha_l}(N'abt\inv) \]
for $N'=-N_{\alpha_{l-2},\alpha_{l-1}-\alpha_{l-2}}N_{{\alpha_{l-1}-\alpha_{l-2}},\alpha_l}$. Since $N_{\alpha,\beta}$ are the structure constants of the Lie algebra, it follows $N=-N'$ from the Jacobi identity.
\end{proof}

Now the same argument as in Lemma~\ref{welldef} allows to prove that $[s_{\alpha_{l-1}}(a),\,x_{\alpha_l}(b)]=1$ for $b\in R_e$, $e\leq0$, what finishes the proof of R3.2$\!\,_{n+1}$.
\end{comment}
\subsection{Computation of the kernel of the map of evaluation at \texorpdfstring{$0$}{0}.}
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.
We also denote by $\overline{H}(\Phi, R, I)$ the subgroup of $\overline{\St}(\Phi, R, I)$ generated by the semisimple root elements $h_\alpha(u)$ and symbols $\{u, v\}$, $u \in (1+I)^\times$, $v \in R^\times$, $\alpha\in \Phi$.

>>>>

\subsection{$\K_2$-groups and symbols} In our computations we use two families of explicit elements of $\K_2(\Phi, R)$ called {\it Steinberg and Dennis--Stein symbols}. Notice that our treatment of symbols follows~\cite{DS73}. Recall that Steinberg symbols are defined for arbitrary $s, t \in R^\times$ as follows:
\begin{equation} \label{eq:steinberg} \{ s, t \}_\alpha = h_\alpha(st) \cdot h_\alpha^{-1}(s) \cdot h_\alpha^{-1}(t). \end{equation}
In turn, Dennis--Stein symbols are defined for arbitrary $a, b\in R$ satisfying $1 + ab \in R^\times$:
\begin{equation} \label{eq:dennis-stein}  \langle a,b \rangle _ \alpha = x_{-\alpha}\left(\tfrac{- b}{1 + ab}\right) \cdot x_{\alpha}(a) \cdot x_{-\alpha}(b) \cdot x_{\alpha}\left(\tfrac{- a}{1+ab}\right) \cdot h_{\alpha}^{-1}(1 + ab). \end{equation} 
Dennis--Stein symbol $\langle a, b \rangle_\alpha$ can be expressed through Steinberg symbols in the special case when either $a$ or $b$ is an invertible element of $R$. More specifically, the following formulae hold (cf.~\cite[p.~250]{DS73}).
\begin{equation} \label{DS-S-relationship} \langle a, b \rangle_\alpha = \{-a, 1+ab\}_\alpha\text{ for } a, 1+ab\in R^\times,\ \
 \{ s, t \}_\alpha = \left\langle -s, \tfrac{1 - t}{s} \right\rangle_\alpha\text{ for } s, t\in R^\times. \end{equation}
%Using~\eqref{eq:conj-h-h} it is easy to verify that \[[h_\alpha(s), h_\beta(t)] = \{s, t^{\langle \alpha, \beta \rangle}\}_\alpha = \{t, s^{\langle \beta, \alpha \rangle}\}_\beta^{-1}.\] 
Steinberg and Dennis--Stein symbols depend only on the length of $\alpha$, in particular they do not depend on $\alpha$ if $\Phi$ is simply-laced. If $\Phi$ happens to be {\it nonsymplectic}, i.\,e. $\Phi \neq \rA_1, \rB_2, \rC_{\geq 3}$, Steinberg symbols are antisymmetric and bimultiplicative, i.\,e. they satisfy the following identities: \begin{equation} \label{eq:symbol-properties} \{ u, st \} = \{ u, s\} \{ u, t \}, \ \{ u, v \} = \{ v, u\}^{-1}. \end{equation}
For these and other properties of symbols we refer the reader to~\cite{DS73}.

@@

\subsection{Unstable Quillen K-groups} \label{sec:quillen}
Throughout this section $G, P, S$ denote three group-valued functors on the category of commutative rings satisfying the following assumptions:
\begin{enumerate} [label=(A\arabic*)]
 \item \label{req:left-exact} $G$ is left exact;
 \item \label{req:subfunc} $P$ is a subfunctor of $G$ such that for all $R$ the group $P(R)$ is a perfect normal subgroup of $G(R)$;
 \item \label{req:uce} There is a natural transformation $\varphi \colon S \to P$ such that $ \varphi \colon S(R) \to P(R)$ is a universal central extension for all $R$;
 \item \label{req:coeq} For every pair $(R, I)$ the functor $S$ has the property that \[S(R \times_{R/I} R) \cong S(R) \times_{S(R/I)} S(R).\]
\end{enumerate}

@@

%TODO: Give the definition of $D_{R, I}$ and of the category of Pairs
\begin{prop}\label{characterization}
For $(G, P, R)$ as above the following statements are true.
\begin{enumerate}
 \item There are natural isomorphisms
  \[ \K_2(G, P, R) \cong \Ker(S(R) \to G(R)), \ \K_3(G, P, R) \cong \HH_3(S(R), \ZZ). \]
 \item The map $\mu \colon S(R, I) \to S(R)$ is the universal relative central extension of the map $S(R) \to S(R/I)$.
       Denote the kernel of $\mu$ by $C(R, I)$.
       Then $C(R, I)$ is naturally isomorphic to the third relative homology group $\HH_3(S(R), S(R/I), \ZZ)$.
 \item There is a natural exact sequence:
 \[ \xymatrix{\K_3(G, P, R) \ar[r] & \K_3(G, P, R/I) \ar@{->>}[r] & C(R, I).} \]
 \end{enumerate}   
\end{prop}
\begin{proof}
 The first claim follows from~\ref{req:subfunc} and~\ref{req:uce} using the standard properties of the plus-construction, see \cite[\S~IV.1]{Kbook}   (cf. Exercises~1.8--1.9 ibid.). %provided one verifies that all the intermediate isomorphisms are natural in $R$.

 Let us verify the second claim. From~\ref{req:left-exact} and $\varphi(G_i) \subseteq \Ker(G(p_i))$ we conclude that $G_0 \cap G_1 \subseteq \Ker(\varphi \colon S(D_{R, I}) \to G(D_{R, I}))$
  therefore $G_0 \cap G_1$ is a central subgroup of $S(D_{R, I})$.
 From~\ref{req:uce} we also obtain that $\HH_1(S(D_{R,I})) = \HH_2(S(D_{R,I})) = 0$ .
 The last two assertions allow us to invoke~\cite[Proposition~6]{Lo78} and conclude that the extension $S(R, I) \to S(R)$ is a universal relative central extension of the coequalizer $\mathrm{coeq}(d_0, d_1)$
  in the sense of~\cite[\S~3]{Lo78}. 
 Notice that this coequalizer coincides with $S(R) \to S(R/I)$ by~\ref{req:coeq}. 
 Finally, the kernel of $\mu$ is isomorphic to $\HH_3(S(R), S(R/I), \ZZ)$ by~\cite[Th{\'e}or{\`e}me~2]{Lo78}.

@@

The principal example of a triple $(G,P,S)$ satisfying the above conditions is as follows.
For $n \geq 5$ take the orthogonal group functor $\mathrm{O}_{2n}(-)$ and its elementary subfunctor $\mathrm{EO}_{2n}(-)$ as functors $G_n$ and $P_n$, respectively. Now set $S_n = \St(\rD_n, -)$. From [Stein's paper] and [Lemma on centrality] it follows that $S_n$ and $P_n$ satisfy~\ref{req:uce}.
 The other requirements can be verified in a straightforward manner. We write $KO_i(n, R)$ as a shorthand for $\K_i(G_n, P_n, R)$.

@@

\section{First injectivity theorem}
If we denote the kernel of the natural transformation $\mu$ from the definition of $\St(\Phi, R, I)$ by $C(\Phi, R, I)$ we obtain a natural exact sequence:
\begin{equation} \label{relSteinberg-def}
 \xymatrix{ 1 \ar[r] & C(\Phi, R, I) \ar[r]^{j} & \St(\Phi, R, I) \ar[r]^{\mu} & \St(\Phi, R) \ar[r]^{\pi} & \St(\Phi, R/I) \ar[r] & 1.}
\end{equation}
The group $C(\Phi, R, I)$ is trivial when $I$ is a splitting ideal of $R$, i.\,e. when the canonical map $R \to R/I$ splits, see \cite[Lemma~8]{S15}.
The following result plays a key role in the proof of~\cref{thm41}.
\begin{lemma} \label{lem:prop41}
In the notation of~\cref{lem:lemma32} assume additionally that the residue field $k$ is of characteristic $\neq 2$.
Then the canonical map $f\colon C(\rD_\ell, B, I) \to C(\rD_\ell, R, I)$ is surjective for $\ell \geq 7$.
\end{lemma}
\begin{proof}
  Consider the following commutative diagram with rows obtained from~\cref{characterization}:
\begin{equation*}\xymatrix{
 KO_3(2\ell, B) \ar[r] \ar[d] & KO_3(2\ell, k[X]) \ar[d]_{f'} \ar@{->>}[r] & \ar[d]_{f} C(\rD_\ell, B, I) \\
 KO_3(2\ell, R) \ar[r]        & KO_3(2\ell, k[X, X^{-1}]) \ar@{->>}[r]        & C(\rD_\ell, R, I).}\end{equation*}
In view of~\cref{lem:Panin-stability} the central vertical map $f'$ is precisely the canonical map $\GW_3^{[0]}(k[X]) \to \GW_3^{[0]}(k[X, X^{-1}])$.
Invoking~\cref{bass-ft} (or~\cite[Corollary~5.3]{Ho05}) we obtain that \[\GW_3^{[0]}(k[X, X^{-1}]) \cong \GW_3^{[0]}(k) \oplus \GW_2^{[3]}(k).\]
By \cite[Lemma~2.2]{FRS12} the group $\GW_2^{[3]}(k)$ is trivial, therefore $f'$ (and hence $f$) are surjective.
\end{proof}

@@

\subsection{Computation of the kernel of the map of evaluation at \texorpdfstring{$0$}{0}.}
For a pair $(R, I)$ we denote by $\overline{\St}(\Phi, R, I)$ the image of the map $\mu$ from~\eqref{relSteinberg-def}, in particular, it is a normal subgroup of $\St(\Phi, R)$.
We also denote by $\St(\Phi, I)$ the subgroup of $\St(\Phi, R)$ generated as a group by root unipotents of level $I$.
It is clear that $\overline{\St}(\Phi, R, I)$ contains $\St(\Phi, I)$ and, in fact, is its normal closure.
We also denote by $\overline{H}(\Phi, R, I)$ the subgroup of $\overline{\St}(\Phi, R, I)$ generated by the semisimple root elements $h_\alpha(u)$ and symbols $\{u, v\}$, $u \in (1+I)^\times$, $v \in R^\times$, $\alpha\in \Phi$.
